---
title: "HW 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lme4)
library(vcd)
library(psych)
```

###1.0
```{r 1_0}
shva = read.table('duryagin_ReductionRussian.txt', header = TRUE)
```

###1.1
```{r 1_1}
shva %>%
  ggplot(aes(x = f2, y = f1, colour = as.factor(vowel))) +
  geom_point() +
  labs(title = 'f2 and f1 of the reduced and stressed vowels', caption = 'Data from Duryagin  2018') + 
  scale_x_reverse() + 
  scale_y_reverse() +
  theme(legend.position = 'none')
```
###1.2.1
```{r 1_2_1}
shva %>%
  ggplot(aes(x = vowel, y = f1, fill=vowel)) +
  geom_boxplot() +
  coord_flip() + 
  labs(title = 'f1 distribution in each vowel', x = element_blank(), caption = 'Data from Duryagin  2018') +
  theme(legend.position = 'none')
```
###1.2.2
```{r 1_2_2}
shva %>%
  ggplot(aes(x = vowel, y = f2, fill = vowel)) +
  geom_boxplot() +
  coord_flip() + 
  labs(title = 'f2 distribution in each vowel', x = element_blank(), caption = 'Data from Duryagin  2018') +
  theme(legend.position = 'none')
```
###1.3
```{r 1_3}
f1_a <- subset(shva$f1, shva$vowel == 'a')
qs <- quantile(f1_a)
IQR_f1_a <- IQR(f1_a) * 1.5
qs2 <- qs[2]
IQR_f1_a_qs2 <- qs2 - IQR_f1_a
qs4 <- qs[4]
IQR_f1_a_qs4 <- IQR_f1_a + qs4

answ <- subset(f1_a, f1_a < IQR_f1_a_qs2 |  f1_a > IQR_f1_a_qs4)
answ
```
###1.4
```{r 1_4}
pc_f1_f2 <- cor(shva$f1, shva$f2, method = 'pearson')
pc_f1_f2
```
###1.5
```{r 1_5}
pc_y <- cor(subset(shva$f1,shva$vowel == 'y'), subset(shva$f2,shva$vowel == 'y'), method = 'pearson')
pc_y
pc_A <- cor(subset(shva$f1,shva$vowel == 'A'), subset(shva$f2,shva$vowel == 'A'), method = 'pearson')
pc_A
pc_a <- cor(subset(shva$f1, shva$vowel == 'a'), subset(shva$f2, shva$vowel == 'a'), method = 'pearson')
pc_a
```
###1.6.1
```{r 1_6_1}
magic_line <- lm(f1 ~ f2, shva)
str(summary(magic_line))
```
###1.6.2
```{r 1_6_2}
r2 <- summary(model)$adj.r.squared
r2
```
###1.6.3
```{r 1_6_3}
shva %>%
  ggplot(aes(x = f2, y = f1, color = as.factor(vowel))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = 'grey') +
  scale_x_reverse() + 
  scale_y_reverse() +
  labs(title='f2 and f1 of the reduced and stressed vowels', caption = 'Data from Duryagin  2018') +
  theme(legend.position = 'none')
```
###1.7.1, 1.7.2
```{r 1_7_1, 1_7_2}
magic_model <- lmer(f1 ~ f2 + (1|vowel), shva)
summary(magic_model)
```

###1.7.3
```{r 1_7_3}
shva %>%
  ggplot(aes(x = f2, y = f1)) +
  geom_point(aes(color = as.factor(vowel))) +
  geom_line(aes(y=fitted(magic_model), color = shva$vowel)) +
  scale_x_reverse() + 
  scale_y_reverse()  +
  labs(title='f2 and f1 of the reduced and stressed vowels', caption = 'Data from Duryagin  2018') + 
  theme(legend.position="none")
```

###2.0
```{r 2_0}
elp <- read.csv("https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/ELP.csv", encoding = "UTF-8")
head(elp)
```

###2.1
```{r 2_1}
attach(elp)
lenth_Sub <- cor.test(Length, SUBTLWF)
lenth_Mean <- cor.test(Length, Mean_RT)
SUB_mean <- cor.test(SUBTLWF, Mean_RT)
lenth_Sub
lenth_Mean
SUB_mean
```
The highest Pearson’s correlaton value have Length and mean_RT.

###2.2
```{r 2_2}
elp %>%
  ggplot(aes(x = SUBTLWF, y = Mean_RT, color = Length)) + 
  geom_point() + 
  scale_color_continuous(low = "lightblue", high = "red")  + 
  facet_wrap( ~ POS) + 
  labs(caption = 'data from (Balota et al. 2007)') +
  scale_x_log10()
```
#### 2.3.1, 2.3.2
```{r 2.3.1, 2.3.2}
very_special <- lm(Mean_RT ~ log(SUBTLWF) + POS, elp)
summary(very_special)
```

###2.3.3
```{r 2_3_3}
elp %>%
  ggplot(aes(x = SUBTLWF, y = Mean_RT, color = Length)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = 'black') +
  scale_color_continuous(low = "lightblue", high = "red")  + 
  scale_x_log10() +
  labs(caption = 'data from (Balota et al. 2007)')

```
###2.4.1, 2.4.2
```{r 2_4_1, 2.4.2}
beauty <- lmer(Mean_RT ~ log(SUBTLWF) + (1 | POS), elp)
summary(beauty)
```
###2.4.3
```{r 2_4_3}
elp %>%
  ggplot(aes(x = log(SUBTLWF), y = Mean_RT, color = POS)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE, colour = 'black') +
  facet_wrap( ~ POS) + 
  labs(x = 'log(SUBTLWF)', caption = 'data from (Balota et al. 2007)') 
  theme(legend.position = 'none')
```

###3.0
```{r 3.0}
d_caus <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv')
```

###3.1
```{r 3.1}
#CrSem, Aux
fisher_Aux_CrSem <- fisher.test(d_caus$CrSem, d_caus$Aux)
fisher_Aux_CrSem
chis_Aux_CrSem <- chisq.test(d_caus$CrSem, d_caus$Aux)
chis_Aux_CrSem

#CeSem, Aux
fisher_Aux_CeSem <- fisher.test(d_caus$CeSem, d_caus$Aux)
fisher_Aux_CeSem
chis_Aux_CeSem <- chisq.test(d_caus$CeSem, d_caus$Aux)
chis_Aux_CeSem

#Aux, CdEvSem
fisher_Aux_CdEvSem <- fisher.test(d_caus$Aux, d_caus$CdEvSem)
fisher_Aux_CdEvSem
chis_Aux_CdEvSem <- chisq.test(d_caus$Aux, d_caus$CdEvSem)
chis_Aux_CdEvSem

#Aux, CeSynt
fisher_Aux_CeSynt <- fisher.test(d_caus$Aux, d_caus$CeSynt)
fisher_Aux_CeSynt
chis_Aux_CeSynt <- chisq.test(d_caus$Aux, d_caus$CeSynt)
chis_Aux_CeSynt

#Aux, EpTrans
fisher_Aux_EPTrans <- fisher.test(d_caus$Aux, d_caus$EPTrans)
fisher_Aux_EPTrans
chis_Aux_EPTrans <- chisq.test(d_caus$Aux, d_caus$EPTrans)
chis_Aux_EPTrans

#Aux, Country
fisher_Aux_Country <- fisher.test(d_caus$Aux, d_caus$Country)
fisher_Aux_Country
chis_Aux_Country <- chisq.test(d_caus$Aux, d_caus$Country)
chis_Aux_Country

#Aux, Domain
fisher_Aux_Domain <- fisher.test(d_caus$Aux, d_caus$Domain)
fisher_Aux_Domain
chis_Aux_Domain <- chisq.test(d_caus$Aux, d_caus$Domain)
chis_Aux_Domain

```
We can`t analyse the correlation between Aux and CeSynt by using Pearson’s Chi-squared Test. The assotiation is sugnificant. It can be seen from the p-value of fisher test.

###3.2
```{r 3.2}
chis_Aux_EPTrans
```
P-value is quite low, so there Aux and EPTrans are not independent.

###3.3
```{r 3.3}
exp_Aux_EPTranse <- chis_Aux_EPTrans[["expected"]]
exp_Aux_EPTranse
```
###3.4
```{r 3.4}
odd_Aux_CrSem <- oddsratio(d_caus$Aux~d_caus$CrSem)
odd_Aux_CrSem
odd_Aux_CeSem <- oddsratio(d_caus$Aux~d_caus$CeSem)
odd_Aux_CeSem
odd_Aux_CdEvSem <- oddsratio(d_caus$Aux~d_caus$CdEvSem)
odd_Aux_CdEvSem
odd_Aux_EPTranse <- oddsratio(d_caus$Aux~d_caus$EPTrans)
odd_Aux_EPTranse
odd_Aux_CeSynt <- oddsratio(d_caus$Aux~d_caus$CeSynt)
odd_Aux_CeSynt
odd_Aux_Country <- oddsratio(d_caus$Aux~d_caus$Country)
odd_Aux_Country
odd_Aux_Domain <- oddsratio(d_caus$Aux~d_caus$Domain)
odd_Aux_Domain
```
###3.5
```{r 3.5}
V_Aux_EPTRanse <- phi(table(d_caus$Aux, d_caus$EPTrans))
V_Aux_EPTRanse
V_Aux_CrSem <- phi(table(d_caus$Aux, d_caus$CrSem))
V_Aux_CrSem
V_Aux_CeSem <- phi(table(d_caus$Aux, d_caus$CeSem))
V_Aux_CeSem
V_Aux_CdEvSem <- phi(table(d_caus$Aux, d_caus$CdEvSem))
V_Aux_CdEvSem
V_Aux_Country <- phi(table(d_caus$Aux, d_caus$Country))
V_Aux_Country
V_Aux_Domain <- phi(table(d_caus$Aux~d_caus$Domain))
V_Aux_Domain
V_Aux_CeSynt <- phi(table(d_caus$Aux~d_caus$CeSynt))
V_Aux_CeSynt
```
For Domain and CeSynt phi cannot be calculated, because there is more than one variable.

###3.6
There is dependency between Aux and EPTranse and Aux and Country.
###3.7
```{r 3.7}
vcd::mosaic(~ Aux + EPTrans + Country, data=d_caus, shade=TRUE, legend=TRUE)
```
###3.8
Firstly because there can be too many pairs to compare. Secondly because we may think that there is no dependency and be wrong.
###3.9
We tested two hyphotheses. H0 - there is no dependency between auxiliary verbs and their transitivity. H1 - there is a dependency between auxiliary verbs and their transitivity.
We find out that the p-value is quite low (p = 0.0001553) so the dependency exists.